---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ggtrack

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/mrjoh3/ggtrack/workflows/R-CMD-check/badge.svg)](https://github.com/mrjoh3/ggtrack/actions)
<!-- badges: end -->

Sometimes in a workplace it is hard to know exactly where or when a specific chart has been produced. This is especially true when someone has copied a chart out of a report to use somewhere else. The `ggtrack` package aims to solve this problem by embedding enough metadata in the charts image to identify the source and the exact time it was produced.

The metadata is added by way of a QR code embedded in a chart "tracking" footer. The QR code can encode any arbitrary text and will append a time stamp. The text is intended to be a URL or a unique id for the original document or chart source.

QR codes can then be scanned using a phone scanner or the [quadrangle package](https://github.com/brianwdavis/quadrangle).

The `ggtrack` footer can also display a corporate logo and additional text.



## Installation

You can install the latest version of project from [Github](https://github.com) with:

``` r
devtools::install_github("mrjoh3/ggtrack")
```

## Development

This package is in the early stages of development. If you have any suggestions, or feature requests please submit an [issue](https://github.com/mrjoh3/ggtrack/issues). All feedback is welcome.

## Example

To start you just need a `ggplot` and some text you wish to encode into the QR. The QR is intended to contain enough information to uniquely identify the report, so a URL, file name or other unique identifier. The QR encode process automatically appends a time stamp. But try to keep the content of the QR code minimal. The for information it is the more pixels its requires and the larger it needs to be. The examples here need a QR code size of 1.8cm to be reliably scanned using a phone off the screen. QR code are encoded using the [qrencoder](https://github.com/hrbrmstr/qrencoder) package.

```{r example, fig.width=6, out.width="80%", fig.align='center', fig.retina=2}
library(ggtrack)
library(ggplot2)

gg <- ggplot(mapping = aes(x = 1:10, y = rnorm(10))) +
  geom_bar(stat = 'identity') +
  theme_minimal()

ggtrack(gg,
        qr_content = 'Report ID: 2c9075a5-4d7e-47a5-8616-55dd88af3dc5')

```

### Add a Logo

A logo can be added either from a local or remote source. For now only `png` and `jpeg/jpg` are supported.

```{r logo, fig.width=6, out.width="80%", fig.align='center', fig.retina=2}
ggtrack(gg,
        qr_content = 'text content here',
        logo = 'https://www.r-project.org/logo/Rlogo.png')

```




### Add a Caption

Captions use the [gridtext](https://github.com/wilkelab/gridtext) package. So you can use both `HTML` and `markdown` to style. Or if you prefer, create your own `grob` and pass that through instead.

```{r caption, fig.width=6, out.width="80%", fig.align='center', fig.retina=2}
ggtrack(gg,
        qr_content = 'For all your image tracking needs: https://github.com/mrjoh3/ggtrack',
        logo = 'https://www.r-project.org/logo/Rlogo.png',
        caption = 'Lots of extra info, or a fancy <span style="color:blue">"grob"</span>.')

```



## Interactive Plots



```{r interactive, fig.width=6, out.width="80%", fig.align='center', fig.retina=2}
ggtrack(gg,
        qr_content = 'For all your image tracking needs: https://github.com/mrjoh3/ggtrack',
        logo = 'https://www.r-project.org/logo/Rlogo.png',
        caption = 'Lots of extra info, or a fancy <span style="color:blue">"grob"</span>.',
        interactive = TRUE)

```



<style>
.hide img {
  float: right;
  margin-bottom: -50%; /* crop position */
}
</style>

<div class="hide">

</div>






