---
title: "ggtrack"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggtrack}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```



## Using ggtrack

To start you just need a `ggplot` and some text you wish to encode into the QR. The QR is intended to contain enough information to uniquely identify the report, so a URL, file name or other unique identifier. The QR encode process automatically appends a time stamp. But try to keep the content of the QR code minimal. The for information it is the more pixels its requires and the larger it needs to be. The examples here need a QR code size of 1.8cm to be reliably scanned using a phone off the screen. QR code are encoded using the [qrencoder](https://github.com/hrbrmstr/qrencoder) package.

```{r example, fig.width=6, fig.height=5, out.width="80%", fig.align='center', fig.retina=2}
library(ggtrack)
library(ggplot2)
library(grid)

gg <- ggplot(mapping = aes(x = 1:10, y = rnorm(10))) +
  geom_bar(stat = 'identity') +
  theme_minimal()

ggtrack(gg,
        qr_content = 'Report ID: 2c9075a5-4d7e-47a5-8616-55dd88af3dc5')

```

### Add a Logo

A logo can be added either from a local or remote source. For now only `png` and `jpeg/jpg` are supported.

```{r logo, fig.width=6, fig.height=5, out.width="80%", fig.align='center', fig.retina=2}
ggtrack(gg,
        qr_content = 'text content here',
        logo = 'https://www.r-project.org/logo/Rlogo.png')

```




### Add a Caption

Captions use the [gridtext](https://github.com/wilkelab/gridtext) package. So you can use both `HTML` and `markdown` to style. Or if you prefer, create your own `grob` and pass that through instead.

```{r caption, fig.width=6, fig.height=5, out.width="80%", fig.align='center', fig.retina=2}
g = ggtrack(gg,
        qr_content = 'For all your image tracking needs: https://github.com/mrjoh3/ggtrack',
        logo = 'https://www.r-project.org/logo/Rlogo.png',
        caption = 'Lots of extra info, or a fancy <span style="color:blue">"grob"</span>.')

grid.draw(g)
```





### Add a download link 

The tracking chart can also be saved and made available for download. You can do this manually with `gplotly::ggsave()` and a simple markdown link as below.

```r
ggsave(g, filename = 'ggrack_simple_download.png')
```

```md
[download png](ggrack_simple_download.png)
```

However, `ggtracker` also has a helper function to facilitate. Using `make_download` you can create either a download link or button. Unlike the markdown link which simply opens the file in the browser; `make_download` creates a proper download link. Note, in the example below it is necessary to specify both location to save the file and the download location. This is only necessary in this example due to the way `pkgdown` handles file locations. In most instances you only needs specify the save location. The button can also be styled by `CSS`.

```{r save link, results='asis'}
dl <- make_download(g, 
                    save_file = c('../man', 'figures', 'ggtrack_chart'),
                    download_file = c('../reference', 'figures', 'ggtrack_chart'),
                    type = 'button',
                    date = '',
                    render = FALSE)
```

`r dl`

Something to consider at this point, with a down load button it is possible to display a plot without the additional metadata. Then pass the tracker object to the download link or button for users. This is a simple way to maintain the aesthetics of your report but still include vital metadata.


## Retrieving Metadata

The metadata from the QR code can be read using the scanner on your phone. If you need to programatically extract this data use the [quadrangle package](https://github.com/brianwdavis/quadrangle). Given the way the QR code are generated, you will need to ensure you specify `flop = TRUE`. This essentially flips the image to replicate what a scanner looking at a screen would capture.

With `quadrangle` you can pass in the entire file or just the QR code. The package will detect the location of the QR code before decoding.

```{r scan}
library(quadrangle)

qr <- qr_scan("../man/figures/ggtrack_chart_.png", flop = TRUE, plot = FALSE)

qr$values$value
```



## Interactive Plots

Adding a tracker to `plotly` interactive plots is currently a work in progress. This option is not straight forward as `geom_GeomCustomAnn()` has not yet been implemented in `plotly`.

```{r interactive, fig.width=6, fig.height=5, out.width="80%", fig.align='center', fig.retina=2}
ggtrack(gg,
        qr_content = 'For all your image tracking needs: https://github.com/mrjoh3/ggtrack',
        logo = 'https://www.r-project.org/logo/Rlogo.png',
        caption = 'Lots of extra info, or a fancy <span style="color:blue">"grob"</span>.',
        interactive = TRUE)

```


<style>
.hide img {
  float: right;
  margin-bottom: -50%; /* crop position */
}
</style>

<div class="hide">

</div>




